Modelling Structured Societies: a Multi-relational Approach to Context Permeability

### 2015-06-21 doc/main.tex
results for context switching reviewed

\subsection{Context Switching}
\noindent In this section, we present the results relative to the context switching model. The difference between this model and the previous one is that the agents no longer interact in the multiple networks at the same time (being able to select any neighbour at a given simulation step). In this model, they become active in a single network at a time. Agents can switch to a different network at the end of each step. The switching mechanism uses a probability associated with the current network in which the agent is currently active. With this new idea of swapping contexts, we covered some space left undeveloped in Antunes and colleagues original work \cite{Antunes2007,Antunes2010}.

As we discussed previously in section \ref{sec:models_cs} (see model \ref{model:context_switching}), the switching probability dictates the frequency with which an agent switches \textit{from} the current network after an encounter has been performed. In an abstract manner, this allows us to model how much time agents spend on each network. A further development of this model will be to assign different preferences to different networks for each agent. This can help us model phenomena such as real-world agents that dedicate more or less time consuming content from different social networks. For now, we attribute this probability to the network to restrain the model complexity.

\subsubsection{Exploring the Switching Probability}
\noindent We first look at the influence of the new parameter (switching probability) in the convergence to consensus. We span the switching probability between $0$ and $1$ in intervals of $0.05$. We do this for 2 networks with \textit{k-regular} and \textit{scale-free} topologies. Figure \ref{fig:ctx_cs_2_kregular} shows the results for two \textit{k-regular} networks: one configuration with $k = 10$ and one configuration with $k = 30$. We use these two configurations to observe the relationship between the number of encounters to achieve consensus, the switching probability parameter and the connectivity of the networks. Remember that with $k=30$ the average path length is also lower (see section \ref{sec:overlapping_kreg}).

@@

\noindent We note that with smaller values of $k$, symmetry in the context switching probability is more important (having the same probability in both networks) if the agent switches less from one of them. This is the case because switching less from one network means spending more time in that network. This means that switching more from the other network can be disruptive to a neighbourhood that has already converged to a sub-convention. This can be observed in configurations with \textit{k-regular} networks (figure \ref{fig:ctx_cs_2_10kreg}) but it is especially apparent in \textit{scale-free} networks (figure \ref{fig:ctx_cs_2_sf_d1}).

@@

When we increase the connectivity (and lower the average path length), both in \textit{k-regular} and \textit{scale-free} networks, the switching probability becomes ``irrelevant:'' in the sense that you can use any probability you want and the results in terms of speed of convergence will be the same. The exception is obviously when considering the switching probability with value $0$. In this case, agents are isolated (they are initially equally distributed throughout the networks) and reaching consensus becomes very hard. --Especially hard since agents are active only in one network: being active in all the networks creates the possibly of disconnected graphs in the networks they are not currently active.

One major difference from the previous model is that the agents no longer can select a neighbour for interaction from any network. The selection has to be made from the network they are currently active in. This causes that agents spending different amounts of time in distinct network topologies are influenced differently -- an agent can be very influential in one network, being a very central node in that context, and marginally important in the other one: in a sense its opinion in one network can be more important to overall convergence than in other network,  due to its position in the network.

To check the influence of connectivity in the new switching dynamics we setup two \textit{scale-free} networks; one with $d=1$ and the other with $d=5$. We then performed the 
same span of switching probability similarly to the previous experiments. The results can be seen in figure \ref{fig:ctx_cs_sf_d5_d1}.

@@

We see that in this case it is more important to switch \textit{from} the network with lower average path length and a forest-like composition ($d=1$) with a probability of at least the same value or more than the one of the other network (with $d=5$). The same can be observed when we mix the two topologies (\textit{k-regular} with \textit{scale-free}), as we can see in figure \ref{fig:ctx_cs_kreg10_sfd1}.

It seems that basically switching less from the less connected network is bad in both cases: spending more time in the network with bigger neighbourhoods allows for a stabler convergence. Sub-conventions can emerge in \textit{scale-free} networks with $d=1$ because usually some network regions are isolated by a single node --the root of the sub-tree they belong to.

@@

\subsubsection{Comparison with Context Permeability}
\noindent Finally, to compare the context switching model with the previous model of context permeability, we constructed the following experiment. We froze the switching probabilities in the values $\zeta={0.25,0.5,0.75}$ and varied the number of networks. (We considered these values because they conveniently characterise the probability in low, medium and high switching) We compare the results with the context permeability in terms of average number of encounters to achieve consensus. Note that we have seen that in some cases, due to the distribution of the number of encounters, the average is not an accurate descriptor for the convergence. Nonetheless, normally the configurations with highly variable outcomes also produce an average that is qualitatively distinct from the rest of the configurations (see for example figure \ref{fig:ctx_perm_kreg} in section \ref{sec:ctx_perm_encounters}). As such, we took the average number of encounters during the simulations to compare what happens in the context switching model with the context permeability model.

@@

Figure \ref{fig:ctx_switching_comp_kreg} shows us that for 2 networks the results in terms of number of encounters for the context switching approximate those we got for context permeability. When we increase the number of networks ($n\ge3$) the configurations with higher switching probability lead the results to be more close to what happens in context permeability. This is no surprise since more switching makes agents switch more often between networks and consequently allows them too choose more often from different neighbourhoods. This is almost the same as having a larger neighbourhood to choose from, which is what happened in the context permeability model.

@@

For \textit{scale-free} networks, figure \ref{fig:ctx_switching_comp_sf} shows us similar
results. It seems that switching with a higher probability displays results similar with those of
context permeability. In this case, adding more networks makes so that lowering switching
probabilities displays slightly worse results. This influence is due to the fact that the
neighbourhoods are considerably smaller in \textit{scale-free} networks -- even with values of $d=5$ -- so the impact of low switching was bound to be noticed.

An increasing number of networks seems to always reduce the number of encounters to achieve consensus in both models. This happens due to what we have called \textit{permeability}. Having multiple points of dissemination that permeate between different social networks enhances the overall convergence to consensus. With permeability there are more previously isolated nodes being reached in sparsely connected networks such as scale-free networks.

%*****************************************
%	Conclusion
%******************************************
\section{Conclusions}
\label{sec:conclusion-future}
\noindent In this article, we analysed and discussed our modelling framework for multi-relational models of social spaces. These results show that not only contexts play an important role in the dissemination of consensus in artificial agent societies, but also that simple mechanisms can generate a great deal of complexity, especially in those scenarios where conventions are being co-learned by multiple agents. Adding more networks to our opinion dynamics model leads to outcomes where consensus was achieved not only more often but also faster. In particular, our results show that achieving convergence (i.e.\ total consensus in a network) is not a matter of connectivity. With scale-free networks we managed to achieve consensus more often and quicker than with less clustered networks. This conclusion is especially important for application with real world social phenomena, many times represented with such networks: full dissemination of a phenomenon (say, gossip-dissemination of a new anti-virus software) may prove impossible with one network only, but is very much eased up by the use of concomitant connections. Moreover, some of the techniques we have rehearsed have proven to speed up even more this now-eased up convergence.

>>>>

      % ***************************************************
      % Context Switching
      % ***************************************************

      \subsection{Context Switching}
      \noindent In this section, we present the results relative to the context switching model. The
      difference between this model and the previous one is that the agents no longer interact in
      multiple networks at the same time (being able to select any neighbour at a given simulation
      step). In this model, they become active in a single network at a time. Agents can switch to a
      different network at the end of each step. The switching mechanism uses a probability
      associated with each network. With this new idea of swapping contexts, we covered some space
      left undeveloped in Antunes and colleagues original work \cite{Antunes2007,Antunes2010}.

      As we discussed in section \ref{sec:models_cs} (see model \ref{model:context_switching}), the
      switching probability dictates the frequency with which an agent switches \textit{from} the
      current network \textit{after} an encounter has been performed. In an abstract manner, this
      allows us to model how much time agents spend on each network. A further development of this
      model will be to assign different preferences to different networks for each agent. This can
      help us model phenomena such as real-world agents that dedicate more or less time consuming
      content from different social networks. For now, we attribute this probability to the network
      to restrain the model complexity.

      \subsubsection{Exploring the Switching Probability}
      \noindent We first look at the influence of the new parameter (switching probability) in the
      convergence to consensus. We span the switching probability between $0$ and $1$ in intervals
      of $0.05$. We do this for 2 networks with \textit{k-regular} and \textit{scale-free}
      topologies. Figure \ref{fig:ctx_cs_2_kregular} shows the results for two \textit{k-regular}
      networks: one configuration with $k = 10$ and one configuration with $k = 30$. We use these
      two configurations to study the relationship between the number of encounters, the switching
      probability parameter, and the connectivity of the networks. Remember that with $k=30$ the
      average path length is also lower (section \ref{sec:overlapping_kreg}).

@@

      \noindent For smaller values of $k$, symmetry in the context switching probability (having the
      same probability in both networks) is more important if the agent switches less from one of
      the networks. Switching less from one network means spending more time in that network. This
      means that switching more from the other network can be disruptive to a neighbourhood that has
      already converged to a sub-convention. This can be observed in configurations with
      \textit{k-regular} networks (figure \ref{fig:ctx_cs_2_10kreg}) but it is especially apparent
      in \textit{scale-free} networks (figure \ref{fig:ctx_cs_2_sf_d1}).

@@

      When we increase the connectivity (and lower the average path length), both in
      \textit{k-regular} and \textit{scale-free} networks, the switching probability becomes
      ``irrelevant'' --for any probability value, the speed of convergence will be the same--. The
      exception is obviously when considering the switching probability with value $0$. In this
      case, agents are isolated (they are initially equally distributed throughout the networks) and
      reaching consensus becomes very hard. (Especially because agents are active only in one network
      and this can create disconnected graphs.)

      One major difference from the previous model is that the agents no longer interact with a
      neighbour from any network at any time. The neighbour selection has to be made from the
      network they are currently active in. This causes the agents to spend different amounts of
      time in different networks. An agent can be very influential in one network -- being a very
      central node in that context -- and marginally important in another: in a sense its opinion in
      one network can be more important to overall convergence than in other network, due to its
      position in the network.

      To explore the contribution of connectivity to the new context switching dynamics, we setup
      two \textit{scale-free} networks; one with $d=1$ and the other with $d=5$. We then spanned the
      switching probability in increments of $0.05$. The results can be seen in figure
      \ref{fig:ctx_cs_sf_d5_d1}.

@@

      In this case it is more important (to convergence speed) to switch \textit{from} the network
      with lower average path length and a forest-like composition (and a probability of at least
      the same value or more than the one of the other network with $d=5$). The same can be observed
      when we mix \textit{k-regular} with \textit{scale-free} networks (see figure
      \ref{fig:ctx_cs_kreg10_sfd1}).

      Switching less from the less connected network is bad in both cases: it is possible that
      spending more time in the network with bigger neighbourhoods allows for a stabler
      convergence. Sub-conventions can emerge in \textit{scale-free} networks with $d=1$ because
      usually some network regions are isolated by a single node -- the root of the sub-tree they
      belong to.

@@

      \subsubsection{Comparison with Context Permeability}
      \noindent Finally, to compare the context switching model with the previous model of context
      permeability, we froze the switching probabilities with values $\zeta={0.25,0.5,0.75}$ and
      varied the number of networks. (We considered these values because they conveniently
      characterise the probability in low, medium and high switching.) We compare the results with
      the context permeability in terms of average number of encounters to achieve consensus. Note
      that in some cases, due to the distribution of the number of encounters, the average is not an
      accurate descriptor for the convergence. Nonetheless, configurations with highly variable
      outcomes usually produce an average that is qualitatively distinct from the rest (see figure
      \ref{fig:ctx_perm_kreg} in section \ref{sec:ctx_perm_encounters}). We took the average number
      of encounters during the simulations to compare the context switching model with the context
      permeability model.

@@

      Figure \ref{fig:ctx_switching_comp_kreg} shows that for 2 networks the results in terms of
      number of encounters for the context switching approximate those for context
      permeability. When we increase the number of networks ($n\ge3$), the configurations with
      higher switching probability lead the results to be closer to what happens in context
      permeability. This is no surprise since more switching makes agents switch more often between
      networks and consequently allows them too choose more often from different
      neighbourhoods. This is almost the same as having a larger neighbourhood to choose from, which
      is what happened in the context permeability model.

@@

      For \textit{scale-free} networks, figure \ref{fig:ctx_switching_comp_sf} shows us similar
      results. Switching with a higher probability displays results similar with those of context
      permeability. In this case, adding more networks makes so that lowering switching
      probabilities displays slightly worse results. This influence comes from the neighbourhoods
      being considerably smaller in \textit{scale-free} networks -- even with values of $d=5$ -- so
      the impact of low switching was bound to be noticed.

      An increasing number of networks seems to always reduce the number of encounters to achieve
      consensus in both models. This happens due to what we called \textit{permeability}. Having
      multiple points of dissemination that permeate between different social networks enhances the
      overall convergence to consensus. With permeability there are more previously isolated nodes
      being reached in sparsely connected networks such as scale-free networks.

      % *****************************************
      % Conclusion
      % ******************************************
      \section{Conclusions}
      \label{sec:conclusion-future}
\noindent In this article, we analysed and discussed our modelling framework for multi-relational models of social spaces. These results show that not only contexts play an important role in the dissemination of consensus in artificial agent societies, but also that simple mechanisms can generate a great deal of complexity, especially in those scenarios where conventions are being co-learned by multiple agents. Adding more networks to our opinion dynamics model leads to outcomes where consensus was achieved not only more often but also faster. In particular, our results show that achieving convergence (i.e.\ total consensus in a network) is not a matter of connectivity. With scale-free networks we managed to achieve consensus more often and quicker than with less clustered networks. This conclusion is especially important for application with real world social phenomena, many times represented with such networks: full dissemination of a phenomenon (say, gossip-dissemination of a new anti-virus software) may prove impossible with one network only, but is very much eased up by the use of concomitant connections. Moreover, some of the techniques we have rehearsed have proven to speed up even more this now-eased up convergence.
